### 2024-05-16 thesis/input_files/data_exploratory.tex
May 15th 2024

This section covers the datasets used in our experiments, in addition, attributes and features will be thoroughly described.
Next, we will explain the ETL pipeline that we set up in order to ease the workflow of our comparison studies.
Finally, in order to to bettern understand patters within the data, we carry out an exploratory data analysis.

\section{GEFCom2014}
The first class of datasets we identified were the ones from the GEFCom 2014 competition. The data is freely hosted on Dr. Hong blog \cite{hong2016probabilistic}.
The main reason is that, these dataset are considered an a excellent test case for comparing predictive models between the EF community. Additionally, the scores of the competing models are freely available, this enables us to carry out a clean and transparent comparative study.
The GEFCom 2014 consisted of four tracks: price, load, wind power and solar power forecasting. 
The GEFCom2014 folder contains a zip file for each of the four tracks.
This thesis work is focuesed on providing forecasts for the load and price quantities.
\\
\subsection{GEFCom2014 Price track}
In the price forecasting track, the goal was to predict the electricity price for the next 24 hours of a single zone. 
The data can be found in the GEFCom2014-P\_V2 zip file alongside a set of instructions and the benchmark forecasts.
This data consisted of time series for the locational marginal price, for the zonal load and for the system load. The data covers the time interval ranging from the 1st of January 2011 to the 15th of June 2013, see figure \ref{fig:price_track_fig1}. 
Additionally, as the competition went on, the real observed data of the previous tasks were made available.
The fifteen target days ordered by task number are: 16/06/2013, 17/06/2013, 24/06/2013, 04/07/2013, 09/07/2013, 13/07/2013, 16/07/2013, 18/07/2013, 19/07/2013, 20/07/2013, 24/07/2013, 25/07/2013, 07/12/2013, 08/12/2013. 
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/price_track_fig1.png}
    \caption{Price track}
    \label{fig:price_track_fig1}
\end{figure}
\subsubsection{EDA}
In exploring this data we started off by plotting the time series of the zonal prices alongside with the toal load and zonal load, see figure \ref{fig:price_track_fig1}.
After that, we plotted the scatter plot for inspecting the correlation between the predictors and the dependent variable, figure \ref{fig:gefcom_zonal_price_vs_zonal_load} and figure \ref{fig:gefcom_zonal_price_vs_total_load}. From these visualisation, we can see their goodness for predicting zonal price since there exist a clear correlation between the candidate predictors and the target variable.

@@

\subsection{Load track}
In the load track, contestant were asked to provide one month ahead hourly probabilistic forecasts on a rolling basis for 15 consecutive rounds. To get started, in the first round, organizers provided 69 months of hourly load data (from 01/01/2005 to 30/10/2010) and 117 months of hourly temperature data (from 01/01/2001 to 30/10/2010). 
As for the price track, the true observed data from the previous track were made available as the competition progressed.
\\
The data for load forecasting is contained in the GEFCom2014-L\_V2 zip file. Within this subfolder, we have a txt with the competition instructions and 15 subfolders one for each of the consecutive tasks. For each of those we have the prediction of the competition benchmark model and the train file upon which we will fit our model.
Each data from the various task folders are shifted by one months between others.
\\
In order to compare our model performance with the winning entries of the GEFCom2014 load competition, we will refer to the Provisional\_Leaderboard\_V2 file contained also in the GEFCom2014 Data directory.
% Notice a quick premise, within this xsls file the authors opted for various scores to compare the entries performance. This choice was motivated by the need 
% to give a proportial weight to each task when averaging between them; that is in order to not penalise too much the tasks with higher point spreads. Furthermore, additional factors such as frequency of benchmark outperformance, and robustness and clarity in methodology explaination.
\\
Notice, the pinball scores for the load track are stored inside the subtab L-score-0/L-score-2 of Provisional\_Leaderboard\_V2.
\subsubsection{EDA}
We applied the same kind of EDA to the GEFCom load data. Figure \ref{fig:gefcom_load_historical} depicts the hisorical load while figure \ref{fig:gefcom_w_avg_historical} reports the historical weather temperature. Finally figure \ref{fig:gefcom_load_vs_w_avg} shows the scatter plot between the the load and the weather.
These plots motivate the selection of weather temperature as predictor for the load.

@@

\section{Germany Switzerland Austria data}
Next, we identified various data source for Germany, Switzerland and Austria in order to further
investigate the techniques studied in this thesis work.
Taking inspiration from the GEFCom2014 format, we firstly created a load track for Germany and Switzerland. 
The source of data has been the \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts} website. It provides data for data for electricity power, electricity potential capacity, energy generation, market prices, environmental measurements and also scenario simulations. Notice, this data is retrieved from the European network of transmission system operators for electricity \href{https://www.entsoe.eu/}{ENTSOE} and made available thanks to the \href{https://ise.fraunhofer.de/en.html}{Fraunhofer Institute for Solar Energy Systems ISE}.
\\
In this experiment, the goal is comparing kernel quantile regression to the other state of the art quantile regressors, see section \ref{ch:prob}.
The target dependent variable is the national load while as predictors we took
the weather temperature, the wind speed, the date expressed in terms of the three attributes day, month and hour, a categorical variable for the day of the week and a binary categorical variable accounting for the holiday effect.

@@

For what concers the data from SECURES-Met we first identified the data interesting to us. Next, we formatted all the csv to a coherent format, gave coloumns meaningful names, sliced for the years 2020 (train) and 2021 (test) and finally we merged all single csv together.

@@

\subsubsection{EDA}
Following are reported the EDA plots alongside with insights that can be extracted from them. We choose to focus the following analysis on the swiss data only given that the plots for the german data are pretty similar.
To get started we visualized the historical load data in figure \ref{fig:CH_historical_load_2021}. As expected, electricity load exhibits a seasonal patter with highs in the winter and lows during the summer season.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_historical_load_2021.png}
    \caption{Switzerland historical load}
    \label{fig:CH_historical_load_2021}
\end{figure}
Next we built scatter plots of the independent continous variables against the load as means of visualizing correlation between them.
Figure \ref{fig:CH_load_vs_temperature_2021} plots load against temperatures while figure \ref{fig:CH_load_vs_wind_speed_2021} depicts load against wind speed. From these plots we can clearly see there exist a relationship that ties wind speeds and temperatures to load.

>>>>

This section covers the datasets used in our experiments, in addition, attributes and features will be thoroughly described.
Next, we will explain the extract transform load pipeline (ETL) that we set up in order to ease the workflow of our comparison studies.
Finally, in order to to better understand patters within the data, we carry out an exploratory data analysis.

\section{GEFCom2014}
The first class of datasets we identified were the ones from the GEFCom 2014 competition. The data is freely hosted on Dr. Hong blog \cite{hong2016probabilistic}.
The main reason is that, these dataset are considered an a excellent test case for comparing predictive models between the EF community. Additionally, the scores of the competing models are freely available, this enables us to carry out a clean and transparent comparative study.
The GEFCom 2014 consisted of four tracks: price, load, wind power and solar power forecasting. 
The GEFCom2014 folder contains a zip file for each of the four tracks.
This thesis work is focused on providing forecasts for the load and price quantities.
\\
\subsection{GEFCom2014 Price track}
In the price forecasting track, the goal was to predict the electricity price for the next 24 hours of a single zone. 
The data can be found in the GEFCom2014-P\_V2 zip file alongside a set of instructions and the benchmark forecasts.
This data consisted of time series for the locational marginal price, for the zonal load and for the system load. The data covers the time interval ranging from the 1st of January 2011 to the 15th of June 2013, see figure \ref{fig:price_track_fig1}. 
Additionally, as the competition went on, the real observed data of the previous tasks were made available.
The fifteen target days ordered by task number are: 16/06/2013, 17/06/2013, 24/06/2013, 04/07/2013, 09/07/2013, 13/07/2013, 16/07/2013, 18/07/2013, 19/07/2013, 20/07/2013, 24/07/2013, 25/07/2013, 07/12/2013, 08/12/2013. 
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/price_track_fig1.png}
    \caption{Price track}
    \label{fig:price_track_fig1}
\end{figure}
\subsubsection{EDA}
In exploring this data we started off by plotting the time series of the zonal prices alongside with the total load and zonal load, see figure \ref{fig:price_track_fig1}.
After that, we plotted the scatter plot for inspecting the correlation between the predictors and the dependent variable, figure \ref{fig:gefcom_zonal_price_vs_zonal_load} and figure \ref{fig:gefcom_zonal_price_vs_total_load}. From these visualisation, we can see their goodness for predicting zonal price since there exists a clear correlation between the candidate predictors and the target variable.

@@

\subsection{Load track}
In the load track, contestant were asked to provide one month ahead hourly probabilistic forecasts on a rolling basis for 15 consecutive rounds. To get started, in the first round, organizers provided 69 months of hourly load data (from 01/01/2005 to 30/10/2010) and 117 months of hourly temperature data (from 01/01/2001 to 30/10/2010). 
As for the price track, the true observed data from the previous track were made available as the competition progressed.
\\
The data for load forecasting is contained in the GEFCom2014-L\_V2 zip file. Within this subfolder, we have a txt file with the competition instructions and 15 subfolders one for each of the consecutive tasks. For each of those we have the prediction of the competition benchmark model and the train file upon which we will fit our model.
Each data from the various task folders are shifted by one months between others.
\\
In order to compare our model performance with the winning entries of the GEFCom2014 load competition, we will refer to the Provisional\_Leaderboard\_V2.csv file contained also in the GEFCom2014 Data directory.
% Notice a quick premise, within this xsls file the authors opted for various scores to compare the entries performance. This choice was motivated by the need 
% to give a proportial weight to each task when averaging between them; that is in order to not penalise too much the tasks with higher point spreads. Furthermore, additional factors such as frequency of benchmark outperformance, and robustness and clarity in methodology explaination.
\\
Notice, the pinball scores for the load track are stored inside the subtab L-score-0/L-score-2 of Provisional\_Leaderboard\_V2.
\subsubsection{EDA}
We applied the same kind of EDA to the GEFCom load data. Figure \ref{fig:gefcom_load_historical} depicts the hisorical load while figure \ref{fig:gefcom_w_avg_historical} reports the historical weather temperature. Finally, figure \ref{fig:gefcom_load_vs_w_avg} shows the scatter plot between the the load and the weather.
These plots motivate the selection of weather temperature as predictors for the load.

@@

\section{Germany Switzerland Austria data}
Next, we identified various data sources for Germany, Switzerland and Austria in order to further
investigate the techniques studied in this thesis work.
Taking inspiration from the GEFCom2014 format, we firstly created a load track for Germany and Switzerland. 
The source of data has been the \href{https://www.energy-charts.info/index.html?l=en&c=DE}{energy charts} website. It provides data for electricity power, electricity potential capacity, energy generation, market prices, environmental measurements and also scenario simulations. Notice, this data is retrieved from the European network of transmission system operators for electricity \href{https://www.entsoe.eu/}{ENTSOE} and made available thanks to the \href{https://ise.fraunhofer.de/en.html}{Fraunhofer Institute for Solar Energy Systems ISE}.
\\
In this experiment, the goal is comparing kernel quantile regression to the other state of the art quantile regressors, see section \ref{ch:prob}.
The target dependent variable is the national load while as predictors we took
the weather temperature, the wind speed, the date expressed in terms of the three attributes day, month and hour, a categorical variable for the day of the week and a binary categorical variable accounting for the holiday effect.

@@

For what concers the data from SECURES-Met we first identified the data interesting to us. Next, we formatted all the csv to a coherent format, gave coloumns meaningful names, sliced for the years 2020 (train) and 2021 (test) and finally we merged all single csv\textsubscript{s} together.

@@

\subsubsection{EDA}
Following are reported the EDA plots alongside with insights that can be extracted from them. We choose to focus the following analysis on the Swiss data only given that the plots for the German data are pretty similar.
To get started we visualized the historical load data in figure \ref{fig:CH_historical_load_2021}. As expected, electricity load exhibits a seasonal patter with highs in the winter and lows during the summer season.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/CH_historical_load_2021.png}
    \caption{Switzerland historical load}
    \label{fig:CH_historical_load_2021}
\end{figure}
Next we built scatter plots of the independent continous variables against the load as means of visualizing correlation between them.
Figure \ref{fig:CH_load_vs_temperature_2021} plots load against temperatures while figure \ref{fig:CH_load_vs_wind_speed_2021} depicts load against wind speed. From these plots we can clearly see there exist a relationship that ties wind speeds and temperatures to load.

### 2024-05-16 thesis/input_files/energymarket.tex
May 15th 2024

\section{Electricity markets}
Electricity markets have gone rapid changes all over the world during the last three decades.
Before this revolution, the power sector was a natural monopoly. In particular, it was characterized by a high vertical integration and by very little market competition. Better technologies in both transmission, generation and distribution are the reasons for such liberal shift. The rationale is simple, a competitive market promotes efficiency and stimulates technical innovation more than a monopoly does.
\\
During the nineties various wholesale electricity markets were established; for example in Chile, Great Britain, the scandinivian block, Australia, New Zealand, the US and Canada. Consequently, new participants, each with a specific task, entered the power market scenario.
\begin{itemize}
    \item Electricity generators produce electricity to be consumed.
    \item Electricity transmission owners 
    move high voltage electricity to electric utility companies whilst ensuring safe and reliable supply.
    \item Electricity grid operators are responsible for scheduling electricity over the transmission network in order to ensure supply demand balance.
    \item Electric utilities deliver power over local lines to consumers. 
    \item Retail energy suppliers purchase electricity in the wholesale market from electricity generators and resell it to consumers.
\end{itemize}
Notice, these functions are not mutually exclusive; that is an entity can provide one or more of these services.

\subsection{The marketplace}
We can differentiate between two kinds of electricity markets: power pools and power exchanges. In power pools, generators bid the prices at which they are willing to produce at different volumes. Then, the market clearing price (MCP) is determined by intersecting the aggregated supply curve and the estimated demand, left panel figure \ref{fig:pool_vs_echange}. Power pools are created on public initiatives of governments. Conversely, power exchanges are created through private agreements between generators, distributors and traders. This is the model followed by most of european countries.
The MCP in power exchanges is determined by the intersection of the aggregate supply curve and the aggregate demand curve, right panel figure \ref{fig:pool_vs_echange}.
\\
It is worth to mention the two biggest power exchanges: Epex and NordPool. Epex operates throughout continental europe. NordPool covers the nordinc and baltic contries.
\\
Moreover, we can differentiate between two popular types of auctions: uniform-price/marginal and pay as bid/discriminatory. Within the uniform price setting, suppliers offering less than the clearing price are paid that price. Analogously, consumers bidding more than the clearing price pay that price. On the other hand, in pay as bid auctions, suppliers are paid the exact price they bid for.

@@

% outages=interruzione
\subsection{Negative prices}
It is not unusual to observe negative prices in electricity markets, even though they are rare.
The causes of negative prices are inflexible power generation plants and low demand. With inflexibility, it is mean the fact that power sources cannot be switched off and restarted quickly and efficiently.
Thus, producers are faced with the decision of either stopping and restarting their power plant or selling their energy for a negative price; that is they pay consumers for consuming their energy.

\subsection{Nodal and zonal pricing}
Grid configuration and physical limits on electricity lines may lead to congestion. Local marginal price (LMP) and zonal market clearing pricing (ZMCP) are two types of pricing schemes utilized to handle it. The former is adopted in EU contries while the latter is used in the US, figure \ref{fig:eu_zonal} and figure \ref{fig:us_local}.
With zonal market clearing pricing, prices may differ between zones but are the same within the same area.
On the other hand, local marginal price is made up by summing the transmission congestion cost, generation marginal cost and the cost of marginal losses at different buses; buses is where a electicity line or several lines are connected.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/eu_zonal.png}
    \caption{Zonal pricing EU \cite{eu_zonal}}
    \label{fig:eu_zonal}
\end{figure}

@@

% pensodino \subsection{Nodal and Zonal Pricing}
\section{Smart grids}
A smart meter is a digital device recording energy consumption in real time and communicating this data to the supplier and the consumer.
Deployment of smart grids and the digitalisation of the electricity network are intended to increase the energy efficiency of our networks.
The reason is that, smart metering and the electricity network digitalisation  will enable system operators to better monitor the network, plan their investments and manage infrastructure. 
All of this motivates for the need of efficient and robust tools to analyze the vast data resulting from the digitalisation of the power sector.

@@

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/electricity-production-by-source.png}
    \caption{Electricity production by source}    
    \label{fig:electricity_production_by_source}
\end{figure}

>>>>

\section{Electricity markets}
Electricity markets have undergone rapid changes all over the world during the last three decades.
Before this revolution, the power sector was a natural monopoly. In particular, it was characterized by a high vertical integration and by very little market competition. Better technologies in both transmission, generation and distribution are the reasons for such liberal shift. The rationale is simple, a competitive market promotes efficiency and stimulates technical innovation more than a monopoly does.
\\
During the nineties various wholesale electricity markets were established; for example in Chile, Great Britain, the scandinivian block, Australia, New Zealand, the US and Canada. Consequently, new participants, each with a specific task, entered the power market scenario.
\begin{itemize}
    \item Electricity generators produce electricity to be consumed.
    \item Electricity transmission owners 
    move high voltage electricity to electric utility companies whilst ensuring safe and reliable supply.
    \item Electricity grid operators are responsible for scheduling electricity over the transmission network in order to ensure supply demand balance.
    \item Electric utilities deliver power over local lines to consumers. 
    \item Retail energy suppliers purchase electricity in the wholesale market from electricity generators and resell it to consumers.
\end{itemize}
Notice, these functions are not mutually exclusive; that is an entity can provide one or more of these services.

\subsection{The marketplace}
We can differentiate between two kinds of electricity markets: power pools and power exchanges. In power pools, generators bid the prices at which they are willing to produce at different volumes. Then, the market clearing price (MCP) is determined by intersecting the aggregated supply curve and the estimated demand, left panel figure \ref{fig:pool_vs_echange}. Power pools are created on public initiatives of governments. Conversely, power exchanges are created through private agreements between generators, distributors and traders. This is the model followed by most of european countries.
The MCP in power exchanges is determined by the intersection of the aggregate supply curve and the aggregate demand curve, right panel figure \ref{fig:pool_vs_echange}.
\\
It is worth to mention the two biggest power exchanges: Epex and NordPool. Epex operates throughout continental europe. NordPool covers the nordinc and baltic contries.
\\
Moreover, we can differentiate between two popular types of auctions: uniform-price/marginal and pay as bid/discriminatory. Within the uniform price setting, suppliers offering for less than the clearing price are paid that price. Analogously, consumers bidding more than the clearing price pay that price. On the other hand, in pay as bid auctions, suppliers are paid the exact price they bid for.

@@

% outages=interruzione
\subsection{Negative prices}
It is not unusual to observe negative prices in electricity markets, even though they are rare.
The causes of negative prices are inflexible power generation plants and low demand. With inflexibility, it is mean the fact that power sources cannot be switched off and restarted quickly and efficiently.
Thus, producers are faced with the decision of either stopping and then restarting their power plant or selling their energy for a negative price; that is they pay consumers for consuming their energy.

\subsection{Nodal and zonal pricing}
Grid configuration and physical limits on electricity lines may lead to congestion. Zonal market clearing pricing (ZMCP) and local marginal price (LMP) are two types of pricing schemes utilized to handle it. The former is adopted in EU contries while the latter is used in the US, figure \ref{fig:eu_zonal} and figure \ref{fig:us_local}.
With zonal market clearing pricing, prices may differ between zones but are the same within the same area.
On the other hand, local marginal price is made up by summing the transmission congestion cost, generation marginal cost and the cost of marginal losses at different buses; buses is where a electicity line or several lines are connected.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/eu_zonal.png}
    \caption{Zonal pricing EU \cite{eu_zonal}}
    \label{fig:eu_zonal}
\end{figure}

@@

% pensodino \subsection{Nodal and Zonal Pricing}
\section{Smart grids}
A smart meter is a digital device recording energy consumption in real time and communicating this data to the supplier and the consumer.
Deployment of smart grids and the digitalisation of the electricity network are intended to increase the energy efficiency of our networks.
The reason is that, smart metering and the electricity network digitalisation  will enable system operators to better monitor the network, plan their investments and manage infrastructure. 
All of this motivates for the need of efficient and robust tools to analyze the vast data resulting from the digitalisation of the power sector see \cite{declean,andreadou2016telecommunication}.

@@

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/electricity-production-by-source.png}
    \caption{Electricity production by source \cite{energy2023statistical}}    
    \label{fig:electricity_production_by_source}
\end{figure}

### 2024-05-16 thesis/input_files/implementation.tex
May 15th 2024

This section is inteneded to explain and aid for reproducibilty studies. Hereafter the specific libraries used and the custom implementatios are thoroughly documented.
\\
For the list of python packages needed see the requirement.txt file.

@@

\subsection{K-nearest neighbours}
The object KNeighborsRegressor of the sklearn module neighbors has been used.

\subsection{Support vector regression}
The object SVR of the sklearn module svm has been used by specifying the liner kernel.

@@

\subsection{Kernel ridge regression}
The object KernelRidge of the sklearn module kernel\_ridge has been used.

\subsection{Kernel support vector regression}
The object SVR of the sklearn module svm has been used by specifying rbf as the kernel parameter.

\section{Probabilistic forecasting}
\subsection{Linear quantile regression}
The implementation of statsmodels.regression.quantile\_regression.QuantReg has been used.
The model is fitted through iterative reweighted least squares.
\subsection{Quantile gradient boosting machine}
The implementation of sklearn.ensemble.GradientBoostingRegressor has been used.
\subsection{Quantile forest}
The implementation of quantile\_forest.RandomForestQuantileRegressor has been used. This estimator is compatible with scikit-learn.
\subsection{Kernel quantile regression}
Kernel quantile regression had no previously implemented open source library, thus the need of implementing our own version.
\\
The scikit-learn team provides a project template for the creation of estimators compatible with scikit-learn functionalities. Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
Our KQR class is initialized by providing a quantile, the regularization term C and the rbf kernel bandwidth gamma; the latter are the two hyperparameters of our custom estimator.
\\
In the fit method, we set up and solve the convex optimisation problem by using the cvxopt library. 
When using this library, it is important to keep two things in mind. First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the Q matrix with no 0.5 in front.
Secondly, in order to specify multiple inequalities we have to stack them an provide as a unique matrix.
\\
Once a solution to the convex problem has been found, we create a mask for the support vectors of the estimator in order to estimate the constant term of our kernel quantile regressor.
\\
In the predict method, we pass a matrix X\_eval of independent variables, next we compute the kernel matrix between X\_train and X\_eval and obtain y\_eval with the formula $y=\alpha^\intercal K+b$.
\\
This estimator is compatible with useful scikit-learn in built methods like gridsearch, crossvalidation and scoring rules. Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}, functions supported by our kernel quantile regression include: gaussian rbf, laplacian, mater 1/2, matern 3/2, matern 5/2, linear, cosine, sigmoid, periodic, polynomial and custom composition of kernels.
% NOTICE, these kernel implementations differ in the form of passed parameters in the former we pass in the numerator gamma while in the latter we pass l, which stands for length scale in the denominator. To cross validate overt the same range of parameter across the different kernels we made sura that gamma=1/l.
\\
Up to now, there are only two open source implementation of the quantile kernel regression. Nevertheless they are both in R, that is there exists no python, matlab or julia open source implementation. Following are reported the results of a comparative study between our own implementation and the one of the R library kernlab
% (i guess it is in C or C++ and then binded to R).
% \subsubsection{Python versus R implementation}
% In this section, a comparison study has been carried out in order to inspect the competitiveness of our implementation with the existing one for the R programming language.

>>>>

This section is inteneded to explain and aid for reproducibilty studies. Hereafter the specific libraries used and the custom implementatios are thoroughly documented.
\\
For the list of python packages needed see the requirement.txt file on \url{https://github.com/luca-pernigo/ThesisKernelMethods}\label{github_repo}.

@@

\subsection{K-nearest neighbours}
The object KNeighborsRegressor of the \href{https://scikit-learn.org/stable/}{sklearn} module neighbors has been used.

\subsection{Support vector regression}
The object SVR of the \href{https://scikit-learn.org/stable/}{sklearn} module svm has been used by specifying the liner kernel.

@@

\subsection{Kernel ridge regression}
The object KernelRidge of the \href{https://scikit-learn.org/stable/}{sklearn} module kernel\_ridge has been used.

\subsection{Kernel support vector regression}
The object SVR of the \href{https://scikit-learn.org/stable/}{sklearn} module svm has been used by specifying rbf as the kernel parameter.

\section{Probabilistic forecasting}
\subsection{Linear quantile regression}
The implementation of \href{https://www.statsmodels.org/dev/generated/statsmodels.regression.quantile_regression.QuantReg.html}{quantile\_regression.QuantReg} from the regression module of statsmodels has been used.
The model is fitted through iterative reweighted least squares.
\subsection{Quantile gradient boosting machine}
The implementation of \href{https://scikit-learn.org/stable/modules/generated/sklearn.ensemble.GradientBoostingRegressor.html}{GradientBoostingRegressor} from the sklearn.ensemble submodule has been used.
\subsection{Quantile forest}
The implementation of \href{https://pypi.org/project/quantile-forest/}{quantile\_forest.RandomForestQuantileRegressor} has been used. This estimator is compatible with scikit-learn API \cite{Johnson2024}.
\subsection{Kernel quantile regression}
Kernel quantile regression had no previously implemented open source library, thus the need of implementing our own version.
\\
The scikit-learn team provides a project template for the creation of estimators compatible with scikit-learn functionalities. Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
Our KQR class is initialized by providing a quantile, the regularization term C and the rbf kernel bandwidth gamma; the latter are the two hyperparameters of our custom estimator.
\\
In the fit method, we set up and solve the convex optimisation problem by using the cvxopt library. 
When using this library, it is important to keep two things in mind. First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the Q matrix with no 0.5 in front.
Secondly, in order to specify multiple inequalities we have to stack them and provide them as a unique matrix.
\\
Once a solution to the convex problem has been found, we create a mask for the support vectors of the estimator in order to estimate the constant term of our kernel quantile regressor.
\\
In the predict method, we pass a matrix $X\_eval$ of independent variables, next we compute the kernel matrix between $X\_train$ and $X\_eval$ and obtain $y\_eval$ with the formula $y=\alpha^\intercal K+b$.
\\
This estimator is compatible with useful scikit-learn in built methods like gridsearch, crossvalidation and scoring rules. Moreover, this code is compatible with sklearn in built kernel functions \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.metrics.pairwise}{sklearn.metrics.pairwise} and \href{https://scikit-learn.org/stable/modules/classes.html#module-sklearn.gaussian_process.kernels}{sklearn.gaussian\_process.kernels}, functions supported by our kernel quantile regression include: gaussian rbf, absolute laplacian, mater 1/2, matern 3/2, matern 5/2, linear, cosine, sigmoid, periodic, polynomial and custom composition of kernels.
% NOTICE, these kernel implementations differ in the form of passed parameters in the former we pass in the numerator gamma while in the latter we pass l, which stands for length scale in the denominator. To cross validate overt the same range of parameter across the different kernels we made sura that gamma=1/l.
\\
Up to now, there are only two open source implementation of the quantile kernel regression. Nevertheless they are both in R, that is there exists no python, matlab or julia open source implementation.

% Following are reported the results of a comparative study between our own implementation and the one of the R library kernlab

% (i guess it is in C or C++ and then binded to R).
% \subsubsection{Python versus R implementation}
% In this section, a comparison study has been carried out in order to inspect the competitiveness of our implementation with the existing one for the R programming language.
